Seismic Hazard Estimation of Northern Iran Using Smoothed Seismicity

### 2016-06-24 abstract.tex
starting again on new completeness adjustments

% 
This article presents a seismic hazard assessment for northern Iran, where a smoothed seismicity approach has been used in combination with an updated seismic catalog and a ground motion prediction equation recently found to yield the best fit with data. We evaluate the hazard over a geographical area including the seismic zones of Azerbaijan, the Alborz Mountain Range, and Kopeh-Dagh; and for completeness, we also consider the contribution of parts of other neighboring seismic zones that fall within our region of interest. In the chosen approach, seismic events are not assigned to specific faults but are instead assumed to be potential seismogenic sources, and are spatially distributed within regular grid cells of size equal to 0.1\textdegree{} in both latitude and longitude. After performing the corresponding magnitude conversions, we decluster both historical and instrumental seismicity catalogs to obtain earthquake rates based on the number of events within each cell, and smooth the results to account for the uncertainty in the spatial distribution of future earthquakes. Seismicity parameters are calculated for each seismic zone separately, and for comparison purposes we also consider the combination of all zones into a single seismotectonic region. Seismic hazard curves are obtained for two models defined based on the assumption that damage occurs above earthquake magnitude thresholds of $M_w$ greater than 4.5 and 5.0, depending on the type of construction. We 
also obtain results for a combined scenario using these two models. The results are presented in terms of expected peak ground acceleration curves and hazard maps, considering exceedance probabilities of 2 and 10 percent in 50 years for rock site conditions. We compare our results to equivalent estimations available in the literature at selected locations and offer our perspective on the differences observed. In conclusion, we find our results to be helpful in understanding seismic hazard for northern Iran from a regional perspective, but highlight the fact that different methods may need to be combined to obtain more robust estimates at specific areas of interest.

>>>>

% 
This article presents a seismic hazard assessment for northern Iran, where a smoothed seismicity approach has been used in combination with an updated seismic catalog and a ground motion prediction equation recently found to yield good fit with data. We evaluate the hazard over a geographical area including the seismic zones of Azerbaijan, the Alborz Mountain Range, and Kopeh-Dagh; and for completeness, we also consider the contribution of parts of other neighboring seismic zones that fall within our region of interest. In the chosen approach, seismic events are not assigned to specific faults but are instead assumed to be potential seismogenic sources, and are spatially distributed within regular grid cells of size equal to 0.1\textdegree{} in both latitude and longitude. After performing the corresponding magnitude conversions, we decluster both historical and instrumental seismicity catalogs to obtain earthquake rates based on the number of events within each cell, and smooth the results to account for the uncertainty in the spatial distribution of future earthquakes. Seismicity parameters are calculated for each seismic zone separately, and for comparison purposes we also consider the combination of all zones into a single seismotectonic region. Seismic hazard curves are obtained for different models defined based on the assumption that damage occurs above certain earthquake magnitude thresholds, depending on the type of construction. We also obtain results for a combined scenario using different models. The results are presented in terms of expected peak ground acceleration curves and hazard maps, considering exceedance probabilities of 2 and 10 percent in 50 years for rock site conditions. We analyze our results in light of equivalent estimations available in the literature and offer our perspective on the differences observed. In conclusion, we find our results to be helpful in understanding seismic hazard for northern Iran from a regional perspective, but highlight the fact that different methods may need to be combined to obtain more robust estimates at specific areas of interest.

### 2016-06-24 completeness.tex
starting again on new completeness adjustments

An important aspect of seismic catalogs is their completeness. This characteristic is measured in terms of the year of completeness and the completeness magnitude, which are used as parameters in the seismic analysis process. In simple terms, the year of completeness is the point in time at which catalog data exhibits a steady rate of event occurrence or record accumulation. This point is typically identified for different magnitude ranges. A simple approach used by \citet{Frankel1995} and others is to select the year at which a plot of the cumulative number of events with respect to time becomes (or starts to resemble) a straight line.

On the other hand, the completeness magnitude---denoted as $M_c$---is the minimum magnitude at which the seismic catalog of a region can be considered complete within a certain time period. While the year of completeness can be determined by inspection of the data, the completeness magnitude often requires a numerical analysis of the catalog. A common approach consists on identifying the magnitude associated with the maximum value of the first derivative of the frequency-magnitude curve. This method is known as the maximum curvature (MAXC) technique. It is widely used and has been implemented in the ZMAP software package for seismic hazard analysis developed by \citet{Wiemer2001}. \citet{Zare2014}, for instance, used this software to identify $M_c$ values for the different regions of the Middle East.

@@

\begin{table*}%[t]
    \centering
    \caption{Year thresholds and magnitude completeness for the seismic zones within the region of interest in this study. Year of completeness is provided for magnitude ranges between 3 and 7, with $\mathrm{\Delta}M = 1$. The completeness magnitude is provided for the sub-catalogs for three year intervals.}
    \begin{tabular}{lccccrcccc}
        \cline{2-10}                                                                                  \\[-1.6ex]
                        & \multicolumn{5}{c}{Year for $M_w$ ranges} 
                                                            & & \multicolumn{3}{c}{$M_c$ for sub-catalogs} \\
        \cline{2-6} \cline{8-10}                                                                      \\[-1.6ex]
                        & 3--4 & 4--5 & 5--6 & 6--7 & \multicolumn{1}{c}{$>7$} 
                                                            & & 1800--1899 & 1900--1999 & 2000--2015  \\[0.6ex]
        \hline                                                                                        \\[-1.6ex]
        Azerbaijan      & 1961 & 1961 & 1900 & 1581 & 1042  & &     6.0    &     5.0    &     4.2     \\
        Alborz          & 2005 & 1961 & 1900 & 1664 &--401  & &     6.2    &     5.0    &     3.5     \\
        Kopeh Dagh      & 2005 & 1950 & 1924 & 1673 &    9  & &     6.1    &     5.0    &     4.0     \\
        Zagros          & 2005 & 1961 & 1950 & 1853 & 1439  & &     5.3    &     4.8    &     3.9     \\
        Central-East    & 2005 & 1961 & 1900 & 1837 &  762  & &     5.5    &     4.8    &     4.0     \\
        Uniform Model   & 2005 & 1961 & 1900 & 1778 &--401  & &     6.1    &     4.8    &     3.8     \\[0.5ex]
        \hline 
    \end{tabular}
    \label{tab:completeness} 
\end{table*}

>>>>

An important aspect of seismic catalogs is their completeness. This characteristic is measured in terms of the year of completeness and the completeness magnitude, which are used as parameters in the seismic analysis process. In simple terms, the year of completeness is the point in time at which catalog data exhibit a steady rate of event occurrence or record accumulation. This point is typically identified for different magnitude ranges independently. A simple approach used by \citet{Frankel1995} and others is to select the year at which a plot of the cumulative number of events with respect to time becomes (or starts to resemble) a straight line.

On the other hand, the completeness magnitude ($M_c$) is the minimum magnitude at which the seismic catalog of a region can be considered complete within a certain time period. While the year of completeness can be determined by inspection of the data, the completeness magnitude often requires a numerical analysis of the given catalog. A common approach consists on identifying the magnitude associated with the maximum value of the first derivative of the frequency-magnitude curve. This method is known as the maximum curvature (MAXC) technique. It is widely used and has been implemented in the ZMAP software package for seismic hazard analysis developed by \citet{Wiemer2001}. \citet{Zare2014}, for instance, used this software to identify $M_c$ values for the different regions of the Middle East.

An alternative approach is the goodness-of-fit test method introduced by \citet{Wiemer2000}.........

@@

\begin{table*}%[t]
    \centering
    \caption{Year thresholds and magnitude completeness for the seismic zones within the region of interest in this study. Year of completeness is provided for magnitude ranges between 3 and 7, with $\mathrm{\Delta}M = 1$. The completeness magnitude is provided for the sub-catalogs for three year intervals.}
    \begin{tabular}{lccccrcccccc}
        \cline{2-12}                                                                                  			\\[-1.6ex]
                        & \multicolumn{5}{c}{Year for $M_w$ ranges} 
                                                            & & \multicolumn{3}{c}{$M_c$ for sub-catalogs} & &$M_c$\\
        \cline{2-6} \cline{8-10}                                                                      			\\[-1.6ex]
                        & 3--4 & 4--5 & 5--6 & 6--7 & \multicolumn{1}{c}{$>7$} 
                                                            & & 1800--1899 & 1900--1999 & 2000--2015  & & Whole	\\[0.6ex]
        \hline                                                                                        			\\[-1.6ex]
        Azerbaijan      & 1961 & 1961 & 1900 & 1581 & 1042  & &     6.0    &     5.0    &     4.2     & &  4.2 	\\
        Alborz          & 2005 & 1961 & 1900 & 1664 &--401  & &     6.2    &     5.0    &     3.5     & &  4.3 	\\
        Kopeh Dagh      & 2005 & 1950 & 1924 & 1673 &    9  & &     6.1    &     5.0    &     4.0     & &  4.9 	\\
        Zagros          & 2005 & 1961 & 1950 & 1853 & 1439  & &     5.3    &     4.8    &     3.9     & &  4.8 	\\
        Central-East    & 2005 & 1961 & 1900 & 1837 &  762  & &     5.5    &     4.8    &     4.0     & &  4.8 	\\
        Uniform Model   & 2005 & 1961 & 1900 & 1778 &--401  & &     6.1    &     4.8    &     3.8     & &  4.8 	\\[0.5ex]
        \hline 
    \end{tabular}
    \label{tab:completeness} 
\end{table*}

### 2016-06-24 composition.tex
starting again on new completeness adjustments

Selection and declustering was also done in consistence with the original catalog, using the method of \citet{Gardner1974}. In summary, we obtained the whole catalog version of \citet{Zare2014}, removed the events from 2000 and forward, complemented the catalog with all $M>3$ events up to 2015 from IIEES, and finally declustered the new complemented catalog removing foreshocks and aftershocks. The resulting set of events compiled for this study is shown in Fig.~\ref{fig:catalog}.

>>>>

Selection and declustering was also done in consistence with the original catalog, using the method of \citet{Gardner1974}. In summary, we obtained the whole catalog version of \citet{Zare2014}, removed the events from 2000 and forward, complemented the catalog with $M>3$ events up to 2015 from IIEES, and finally declustered the new complemented catalog removing foreshocks and aftershocks. The resulting set of events compiled for this study is shown in Fig.~\ref{fig:catalog}, which includes the spatial distribution within the region of interest for both historical and instrumental events; and in Fig.~\ref{fig:scatter}, which presents the distribution of events in time and magnitude for each of the seismic zones, as well as for the complete region of interest (hereafter referred as the uniform model).

\begin{figure*} [t]
	\centering
	\includegraphics[width=\textwidth]{figures/pdf/figure-04}
	\caption{Magnitude-time distribution of earthquakes in each of the seismic zones, and the complete region of interests (here referred to as the uniform model).}
	\label{fig:scatter}
\end{figure*}

### 2016-06-24 introduction.tex
starting again on new completeness adjustments

This article presents a new seismic hazard assessment for the northern portion of the Iranian plateau, using a smoothed seismicity approach in combination with an updated seismic catalog and a ground motion prediction equation recently evaluated and found to be the most reliable attenuation relationship to predict peak ground response in the region.

@@

According to this tectonic setting and the geologic provinces of the plateau, Iranian earthquakes have been categorized into various seismic zones, ranging from the definition of four to nine major seismic zones in the more traditional studies \citep[e.g.,][]{Stocklin1968, Takin1972, Berberian1976}, and up to twenty to twenty-three seismotectonic provinces in the more elaborate ones \citep[e.g.,][]{Nowroozi1976, Tavakoli1999}. Contemporary, well-accepted models divide Iran into five major tectonic regions: Azerbaijan-Alborz, Kopeh-Dagh, Zagros, Central-East Iran, and Makran \citep[e.g.,][]{Mirzaei1998}. More recently, in a study aimed to develop a uniform earthquake catalog for Iran, \citet{Karimiparidari2013} further divided the Azerbaijan-Alborz region into two separate seismic zones: Azerbaijan and the Alborz Mountain Range, hereafter referred to simply as Alborz. We adopt the latter definition of six seismic seismic zones and focus our attention on those most relevant to the northern part of the country, Azerbaijan, Alborz and Kopeh-Dagh.

Over the last three decades there have been multiple efforts to better estimate the seismic parameters and hazard, and improve the assessment of seismic risk in Iran \citep[e.g.,][]{Tavakoli1999, Moinfar_2000_Chap, Ghodrati2003, Moinfar_2012_WCEE, Khodaverdian_2016_BSSA}. The most recent of these studies \citep{Khodaverdian_2016_BSSA} focuses in the estimation of spatially-variable seismic parameters for the whole country using a similar approach to the one we used here. We, however, focus our attention in the northern three seismic zones of the country assuming regionally-based seismic parameters. Altogether, these previous studies have helped improve the Iranian Code of Practice for Seismic Resistant Design of Buildings \citep{BHRC2014}. Such improvements are reflected in lower overall expected damage ratios, as reported by \citet{Ghodrati2013} for the city of Tehran. However, other indicators suggest persistent high levels of seismic risk. Therefore, more complete and accurate assessments of seismic hazard---incorporating the latest available data and alternative methods---are still deemed necessary.

@@

In the following sections we describe in greater detail our region of interest, review the associated tectonics and seismicity, and provide additional information about the adopted approach. Significant aspects of our work include the composition of the earthquake catalog, including catalog completeness, the conversion of earthquake magnitudes, the selection of an appropriate attenuation relationship for the region, and the estimation of the Gutenberg-Richter seismicity parameters. Our analysis includes seismic hazard curves and maps for peak ground acceleration (PGA) considering exceedance probabilities of 2 and 10 percent in 50 years for rock site conditions, considering different magnitude thresholds. We compare our results to equivalent estimations (hazard curves) available in the literature for selected locations and conclude with a discussion in light of other previous hazard studies.

>>>>

This article presents a new seismic hazard assessment for the northern portion of the Iranian plateau, using a smoothed seismicity approach in combination with an updated seismic catalog and a ground motion prediction equation recently evaluated and found to fit well with observations of peak ground response in the region.

@@

According to this tectonic setting and the geologic provinces of the plateau, Iranian earthquakes have been categorized into various seismic zones, ranging from the definition of four to nine major seismic zones in the more traditional studies \citep[e.g.,][]{Stocklin1968, Takin1972, Berberian1976}, and up to twenty to twenty-three seismotectonic provinces in the most elaborate ones \citep[e.g.,][]{Nowroozi1976, Tavakoli1999}. Contemporary, well-accepted models divide Iran into five major tectonic regions: Azerbaijan-Alborz, Kopeh-Dagh, Zagros, Central-East Iran, and Makran \citep[e.g.,][]{Mirzaei1998}. More recently, in a study aimed to develop a uniform earthquake catalog for Iran, \citet{Karimiparidari2013} further divided the Azerbaijan-Alborz region into two separate seismic zones: Azerbaijan and the Alborz Mountain Range, hereafter referred to simply as Alborz. We adopt the latter definition of six seismic seismic zones and focus our attention on those most relevant to the northern part of the country, Azerbaijan, Alborz and Kopeh-Dagh.

Over the last three decades there have been multiple efforts to better estimate the seismic parameters and hazard, and improve the assessment of seismic risk in Iran \citep[e.g.,][]{Tavakoli1999, Moinfar_2000_Chap, Ghodrati2003, Moinfar_2012_WCEE, Khodaverdian_2016_BSSA}. The most recent of these studies \citep{Khodaverdian_2016_BSSA} focuses in the estimation of spatially-variable seismic parameters for the whole country using a similar approach to the one we used here. We, however, focus our attention in the northern three seismic zones of the country assuming regionally-based and smoothed seismic parameters. Altogether, these previous studies have helped improve the Iranian Code of Practice for Seismic Resistant Design of Buildings \citep{BHRC2014}. Such improvements are reflected in lower overall expected damage ratios, as reported by \citet{Ghodrati2013} for the city of Tehran. However, other indicators suggest persistent high levels of seismic risk. Therefore, more complete and accurate assessments of seismic hazard---incorporating the latest available data and alternative methods---are still necessary.

@@

In the following sections we describe in greater detail our region of interest, review the associated tectonics and seismicity, and provide additional information about the adopted approach. Significant aspects of our work include the composition of the earthquake catalog, including catalog completeness, the conversion of earthquake magnitudes, the selection of an appropriate attenuation relationship for the region, and the estimation of the Gutenberg-Richter seismicity parameters. Our analysis includes seismic hazard curves and maps for peak ground acceleration (PGA) considering exceedance probabilities of 2 and 10 percent in 50 years for rock site conditions using different magnitude thresholds. We analyze our results in light of equivalent estimations available in the literature and summarize relevant conclusions.
